Classless Java

### 2015-11-17 paper/sections/RelatedWork.tex
added references. revise related work.

\subsection{Multiple Inheritance in Object oriented languages}
Many authors have argued in favour or against multiple inheritance.
It provides expressive power, but it is difficult to model, and implement and can create programs that are hard to reason about.
These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, and the yo-yo problem~\cite{taenzer1989problems}.
\marco{add citation on the text over.}
\yanlin{added.}
To conciliate the need for expressive power and the need for simplicity, 
lots of models have been proposed in the past few
years, including mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
They provide novel programming architecture models in the
OO paradigm. 
\begin{enumerate}
\item  Mixins allows to name components that can be applied to various classes as
reusable functionality units. However,  they suffer from linearisation: the order of mixin application is relevant in often subtitle and undesired ways. constraints. This hinders their
usability and their ability of resolving conflicts:
 the linearisation (total ordering) of mixin inheritance cannot provide a satisfactory resolution in all cases and restricts the flexibility of mixin composition.

 To fight those limitations, an algebra of mixin operators is introduced~\cite{CMS-elena-zucca-davide-ancona}, but this raised the complexity of the approach, especially when constructors and fields are considered~\cite{Ecoop-FJigsaw-marco}.

\item 
Simplifying the mixin algebras approach, traits draw a strong line between units of reuse (traits) and object factories (classes)
In this model, traits~\cite{Scharli2003} units of reusable code,  containing only methods as
reusable functionalities. Thus, no state/state initialization is considered.

@@

\subsection{Multiple Inheritance in Java}
Since Java 8 default methods are introduced, which enables concrete method
implementation to be defined (via the \textbf{default} keyword) inside
interfaces. Since Java supports implementation of multiple interfaces (instead of extension of a single
class), the introduction of default methods opens the gate for various flavours of multiple
inheritance in Java, using interfaces. 
Former work by Bono.et. al.~\cite{bono14}.
provides details on mimicking traits~\cite{sharly-thesis} through interfaces.

@@

One famous attempt in this direction is provided by 
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of \textbf{this}, changing its meaning along with
inheritance.
However when invoking a method with MyType in a parameter position, the exact type of the receiver must be known.
This is huge limitation in class based object oriented programming,
and is exasperated by interface-only programming as we propose: no type is ever exact since classes are not part of the language.
A recent article~\cite{igarashi Matching MyType to subtyping}
lights up this topic, since
they propose two features: exact statements and non-inheritable methods; both related to our work:
any method generated inside of the \Q@of@ method is indeed non-inheritable, since there is no class name to extends from, and exact statements (a form of wild-card capture on the exact run-time type)
could capture the ``exact type'' of an object even in a class-less environment.

>>>>

\subsection{Multiple Inheritance in Object oriented languages}
Many authors have argued in favour or against multiple inheritance.  It provides
expressive power, but it is difficult to model, and implement and can create
programs that are hard to reason about.  These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods,
and the yo-yo problem~\cite{taenzer1989problems}.  To conciliate the need for
expressive power and the need for simplicity, lots of models have been proposed
in the past few years, including mixins~\cite{bracha90mixin} and
traits~\cite{scharli03traits}.  They provide novel programming architecture
models in the OO paradigm.

\begin{enumerate}
\item Mixins allows to name components that can be applied to various classes as
  reusable functionality units. However, they suffer from linearisation: the
  order of mixin application is relevant in often subtitle and undesired
  ways. constraints. This hinders their usability and their ability of resolving
  conflicts: the linearisation (total ordering) of mixin inheritance cannot
  provide a satisfactory resolution in all cases and restricts the flexibility
  of mixin composition.

  To fight those limitations, an algebra of mixin operators is
  introduced~\cite{ancona2002calculus}, but this raised the complexity of the
  approach, especially when constructors and fields are
  considered~\cite{marco09FJigsaw}.

\item Simplifying the mixin algebras approach, traits draw a strong line between
  units of reuse (traits) and object factories (classes) In this model,
  traits~\cite{scharli03traits} units of reusable code, containing only methods
  as reusable functionalities. Thus, no state/state initialization is
  considered.

@@

\subsection{Multiple Inheritance in Java}
Since Java 8 default methods are introduced, concrete method implementation are
allowed to be defined (via the \textbf{default} keyword) inside
interfaces. Since Java supports implementation of multiple interfaces (instead
of extension of a single class), the introduction of default methods opens the
gate for various flavours of multiple inheritance in Java, using interfaces.
Former work by Bono.et. al.~\cite{bono14}. provides details on mimicking traits
through interfaces.

@@

One famous attempt in this direction is provided by
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
\textbf{this}, changing its meaning along with inheritance.  However when
invoking a method with MyType in a parameter position, the exact type of the
receiver must be known.  This is huge limitation in class based object oriented
programming, and is exasperated by interface-only programming as we propose: no
type is ever exact since classes are not part of the language.  A recent
article~\cite{Saito2013933} lights up this topic, since they propose two
features: exact statements \yanlin{do we support exact statements?} and
non-inheritable methods; both related to our work: any method generated inside
of the \Q@of@ method is indeed non-inheritable, since there is no class name to
extends from, and exact statements (a form of wild-card capture on the exact
run-time type) could capture the ``exact type'' of an object even in a
class-less environment.
